### 2021-05-23 chaps/thesis/chap3/sections/numeric.tex

% !TEX root = ../chap3.tex

\section{Résultats numériques}

\subsection{Calcul de stabilité avec les équations de Maxwell}