### 2021-05-26 chaps/thesis/chap3/sections/numeric.tex
📝 modification of structure of chap3

% !TEX root = ../chap3.tex

\section{Résultats numériques}