### 2021-06-01 chaps/thesis/chap2/chapitre2.tex
💄 insecable space in title

% !TEX root = ../main.tex
\renewcommand{\localPath}{chap2}

\chapter{Modèle hybride linéarisé dans~le~cas~$1dx-1dv$}

%\newcommand{\commentaire}[2][]{\ifthenelse{\equal{#1}{}}{\textbf{#2}}{\ifthenelse{\equal{#1}{Anais}}{\textcolor{magenta}{#2}}{\ifthenelse{\equal{#1}{Nicolas}}{\textcolor{orange}{#2}}{\textcolor{teal}{#2}\footnote{Commentaire rédigé par #1}}}}}

%% section 1
\input{\localPath/sections/intro}
%% section 2
\input{\localPath/sections/models}
%% section 3
\input{\localPath/sections/geom}
%% section 4
\input{\localPath/sections/scheme}
%% section 5
\input{\localPath/sections/dispersion}
%% section 6
\input{\localPath/sections/limit}
%% section 7
\input{\localPath/sections/compare}

% annexe
\begin{subappendices}
\input{\localPath/sections/annexe_rel_disp.tex}
\end{subappendices}