### 2021-06-01 chaps/thesis/chap3/chapitre3.tex

% !TEX root = ../main.tex
\renewcommand{\localPath}{chap3}

\chapter{Modèle hybride linéarisé dans~le~cas~$1dz-3dv$}

%% section 1
\input{\localPath/sections/intro}
%% section 2
\input{\localPath/sections/models}
%% section 3
\input{\localPath/sections/scheme}
%% section 4
\input{\localPath/sections/numeric}
%% section 5
\input{\localPath/sections/approx}
%% section 6
\input{\localPath/sections/numeric_m}
%% section 7
\input{\localPath/sections/info}