### 2021-06-04 chaps/thesis/chap3/sections/models.tex
📝 add filtrage

% !TEX root = ../chap3.tex

\section{Présentation du modèle}

\begin{align}
  \label{eq:VHM:jx}
  &\pdv{j_{c,x}}{t} = \Omega_{pe}^2 { E}_x - { j}_{c,y}B_0,\\
  \label{eq:VHM:jy}
  &\pdv{j_{c,y}}{t} = \Omega_{pe}^2 { E}_y + { j}_{c,x}B_0,\\
  \label{eq:VHM:Bx}
  &\pdv{B_{x  }}{t} = \partial_z E_y,\\
  \label{eq:VHM:By}
  &\pdv{B_{y  }}{t} = -\partial_z E_x,\\
  \label{eq:VHM:Ex}
  &\pdv{E_{x  }}{t} =-\partial_z B_y - {j}_{c,x} + \int v_x f_h \dd{\vb{v}},\\
  \label{eq:VHM:Ey}
  &\pdv{E_{y  }}{t} =\partial_z B_x - {j}_{c,y} + \int v_y f_h \dd{\vb{v}}, \\
  \label{eq:VHM:fh}
  &\pdv{f_{h  }}{t} +  v_z\partial_z f_h - \left(E_x + v_y B_0 - v_zB_y\right)\partial_{v_x} f_h\\
  &\hspace{2cm}- \left(E_y - v_x B_0 + v_z B_x\right)\partial_{v_y} f_h- \left(v_x B_y- v_y B_x\right)\partial_{v_z} f_h = 0.\nonumber
\end{align}