### 2021-08-04 chaps/thesis/intro/introduction.tex
📝 remove useless notes

% !TEX root = ../main.tex
\renewcommand{\localPath}{intro}

\setcounter{chapter}{-1}
\chapter{Introduction}

% section intro
\input{\localPath/sections/intro.tex}
% section modélisation
\input{\localPath/sections/models.tex}
% section schémas
\input{\localPath/sections/scheme.tex}

% annexe
\begin{subappendices}
\input{\localPath/sections/annexe.tex}
\end{subappendices}